---
title: "NH Daily File"
subtitle: "Designating NH Residence using MDS Data - 2020"
author: 
  - "Andrew Tungate"
date: 'June 10th, 2021'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  fig.showtext = TRUE,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(
  base_color = "#23395b"
  )
xaringanExtra::use_tile_view()
```

```{r create_r_objects, include=FALSE, warning=FALSE}

library(here)
library(knitr)
library(kableExtra)
library(ggplot2)

# Fields table
fields = read.csv(here("data", "fields.csv"), check.names = FALSE)

# Example records
records = data.frame(bene_id = c(rep(24601, 27), rep(NA, 3), rep(24601, 10), rep(8675309, 10)),
                     sasdt = c(seq(as.Date("2020-01-01"), as.Date("2020-01-27"), 1L),
                               rep(NA, 3),
                               seq(as.Date("2020-12-22"), as.Date("2020-12-31"), 1L),
                               seq(as.Date("2020-01-01"), as.Date("2020-01-10"), 1L)),
                     sas_month = NA,
                     last_trgt = as.Date(c(rep("2019-10-15", 11), rep("2020-01-12", 16), rep(NA, 3),
                                           rep("2020-09-06", 10), rep("2019-12-22", 10))),
                     entry_sasdt = as.Date(c(rep("2019-03-15", 27), rep(NA, 3), 
                                             rep("2020-06-09", 10), rep("2019-12-22", 6), rep(NA, 4))),
                     in_nh_today = c(rep("Y", 27), rep(NA, 3), rep("Y", 17), rep("N", 3)),
                     episode_length30d = c(seq(78, 104, 1), rep(NA, 3),
                                           seq(431, 440, 1), 10:16, rep(0, 3)),
                     days_since_discharge = c(rep(".", 27), rep(NA, 3), rep(".", 16), 0:3),
                     deceased_yn = c(rep("N", 27), rep(NA, 3), rep("N", 16), "Y", rep("P", 3)))
records$sas_month = toupper(format(records$sasdt, format = "%b%Y"))

temp_cols = c("sasdt", "last_trgt", "entry_sasdt")
records[temp_cols] = lapply(records[temp_cols], function(x) toupper(format(x, format = "%d%b%Y")))
records[is.na(records)] = "..."
records[47:50, "entry_sasdt"] = "."

# Data for churn plot
churn = read.csv(here("data", "churn.csv"))
churn$sas_month = as.Date(paste0("01", churn$sas_month), format = "%d%b-%y")
churn$pct_admitted = churn$stay_lt_30d / churn$N
churn$pct_discharged = churn$discharge_lt_30d / churn$N
churn$pct_dschrgdeath = churn$deceased_yn / churn$discharge_lt_30d

```


## Overview

### Problem: 

--

We needed a way to track when a beneficiary was in a nursing home (NH)

--

### Possible Solutions:

--

 1. A daily file with the membership-status (i.e., whether a beneficiary was in the NH or in the community) over the course of the year

--

 2. A monthly file with the membership-status over the course of the year

--

 3. A finder file with info on if a beneficiary was in a NH at any point during the year

--

 4. The NF Stays file (which Albert will cover in more detail)
 
 
 
---
class: inverse center middle

# Option 1. The NH Daily File


---
layout: true

## 2020 NH Daily File

---

### Characteristics

--

 - Includes all bene's that were in a NH at any point in 2020*

--

 - Includes 366 days for each beneficiary (2020 was a leap year)

--

 - Has between 1 and 2 billion rows

--

 - Based on MDS 3.0 data; not based on claims data

--

 - Only non-MDS field is the `deceased` field


---

### Fields

```{r, echo=FALSE}

kable(fields, format = 'html') %>%
  kable_styling(bootstrap_options = c("condensed", "striped"), 
                full_width = FALSE, font_size = 12)

```

---

### Record Layout
```{r, echo=FALSE}

kable(records, format = 'html') %>%
  kable_styling(bootstrap_options = c("condensed", "striped"), 
                full_width = FALSE, font_size = 12) %>%
  scroll_box(height = "65%")

```

---
layout: true

## Examples 

---

### % of NH Residents Discharged & Expired in the Same Month

```{r, echo = FALSE}

churn %>%
  ggplot(aes(x = sas_month, y = pct_dschrgdeath)) +
  geom_line(stat = "identity", size = 1.5, color = "#23395b") +
  geom_point(stat = "identity", shape = 21, color = "#23395b", fill = "white", size = 3) +
  scale_y_continuous(labels = scales::label_percent(1L), limits = c(0, .5)) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %y") +
  theme_minimal() +
  xaringanthemer::theme_xaringan(text_font_size = 12) + 
  theme(panel.grid.minor.x = element_blank(),
        axis.title = element_blank())


```



---

### Point-in-time checks

--

 - Compare to other sources (e.g., [The MDS 3.0 Frequency Report](https://www.cms.gov/Research-Statistics-Data-and-Systems/Computer-Data-and-Systems/Minimum-Data-Set-3-0-Public-Reports/Minimum-Data-Set-3-0-Frequency-Report)) by choosing a point-in-time

--

```{sql, eval = FALSE}
proc sql; 
  CREATE TABLE  eoy_base AS
  SELECT        *
  FROM          daily_file_name
  WHERE         sasdt = "31DEC2020"d    #<<
                AND in_nh_today = "Y";
quit;
```

---

### Point-in-time checks

 - Compare to other sources (e.g., [The MDS 3.0 Frequency Report](https://www.cms.gov/Research-Statistics-Data-and-Systems/Computer-Data-and-Systems/Minimum-Data-Set-3-0-Public-Reports/Minimum-Data-Set-3-0-Frequency-Report)) by choosing a point-in-time

```{sql, eval = FALSE}
proc sql; 
  CREATE TABLE  eom_base AS
  SELECT        *
  FROM          daily_file_name
  WHERE         in_nh_today = "Y"
  GROUP BY      bene_id, sas_month     #<<
  HAVING        sasdt = max(sasdt);    #<<
quit;
```


---

### Drawbacks

--

 - The file is...pretty big
 
--

 - Some of the error from the MDS assessments will carry forward
 
--
 
 - There would need to be separate files for separate years

--

 - Although it has been QA'd, formal tests have not be run on it (yet)


---
layout: false



## Outstanding Methodological Questions

--

 - Which variables should be included? 
  - Should it be pure MDS?
  
--
  
 - Do we want to retain the 150 & 30 day rules?
 
--

 - When `deceased = "P"`, either:
  - Remove the rows
  - Select fields that should be blank

--

 - Odds and ends:
  - Finalize variable names
  - SAS formats

  


---

## Future Directions

--

 - Figure out the use-cases

--

 - Create yearly files for past years (e.g., 2018 NH Daily File, 2019 NH Daily file)
 
--
 
 - Create a monthly file for the current year
 
--
 
 - Make script portable for derivation in the CCW
 
--
 
 - Create tests & documentation


---
class: center, middle

## Any Questions?



Slides created via the R packages:
[**xaringan**](https://github.com/yihui/xaringan)<br>
[gadenbuie/xaringanthemer](https://github.com/gadenbuie/xaringanthemer)

*Note: Need to find a slide? Hit "o" on your keyboard*


